### 2009-11-16 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

\chapter{Netbeams in Depth: Requirements Analysis and Specification}

This chapter presents NetBEAMS and its implementation details, showing its
main application of the the environmental sensor network deployed at the
Romberg Tiburon Center (RTC) as it is the main case study.

@@

\section{DSP Message and DSP Messages Container}

>>>>

\chapter{Netbeams in Depth: Requirements Analysis and Specification}
\label{chap:netbeams-overview}
This chapter presents NetBEAMS and its implementation details, showing its
main application of the the environmental sensor network deployed at the
Romberg Tiburon Center (RTC) as it is the main case study.

@@

\section{DSP Message and DSP Messages Container}
\label{sec:dsp-message}

### 2009-11-16 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

\section{Analysis of the Location of Sensor Data Taxonomy}

@@

@@

@@

Other projects have been using the XML data models. It falls into the same
category as the Relational Model, since the XML documents, when in the context
of a database, must be complaint to an XML Schema. As shown in section 3, this
model uses XPath technology for querying documents, although some hibrid
technologies may still use SQL for that matter. In this way, an instance of
such schema dependent types can be seen in figure
\ref{fig:persistence-example-relational}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.65]{../diagrams/persistence-example-relational}
  \caption{Instance of a Relational Model Database Prototype}
  \label{fig:persistence-example-relational}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/KVP-on-Relational-Model}
  \caption{KVP Data Model implementation using Relational Model}
  \label{fig:KVP-on-Relational-Model}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/persistence-example-relational-kvp}
  \caption{KVP over Relational Model Database Prototype}
  \label{fig:persistence-example-relational-kvp}
\end{figure}

@@

Sensor Network data can be saved in either Centralized or Distributed Systems.
While Centralized Database Systems are easier to manage, it may face challenges
regarding its data. For this reason, the use of partitioning or sharding would
be an addition to persist sensor data.

>>>>

\section{Analysis of the Location of Sensor Data Taxonomy}
\label{sec:sn-data-location}

@@

In order to support this strategy, a technique called Database Sharding
\cite{db-shard02} is a technique used in distributed database systems that
uses denormalized data models to improve scalability in terms of data storage
and data location. It is a type of database system patitioning where
de-normalized data is spread across different other hosts, or shard \cite{db-shard01}.
Different models of database sharding can be developed \cite{db-shard-schemas}.

\begin{itemize}
  \item Vertical Partitioning: each feature of the system, usually represented
  by a table in the relational model, is placed in a different server host. For
  example, each sensor in a given network can be divided by its device type;
  \item Range Based Partitioning: after partitioning each feature on a
  different server host, different multiple shards, or similar hosts, can be
  added by using the range of a given key of the entity of a function. For
  example, all sensor types in a geographic location may contain its own shard
  based on its IP address;
  \item Key or Hash Based Partitioning: a given value for a given property of
  an instance of data that is chosen to be used in a hash function, whose
  output determines which shard to use;
  \item Directory Based Partitioning: a simple lookup service is used to
  identify which database shard to use, where the lookup table can be stored
  in a secondary area. Each approach for database sharding used be used
  together on a given function depending on the needs for performance.
\end{itemize}

For example, figure \ref{fig:database-sharding-by-region} is an example of
data storage with shards \cite{db-shard01} related the location of data, where
each shard is assumed to be placed in different server host. Since the data
model is denormalized, the organization of data is entirely placed in each of
these shards. As it is shown, the shards by location can be defined by a shard
key, as described before. The horizontal scalability is given related to
positions, while the vertical by the shards added.

\begin{figure}
  \centering
  \includegraphics[scale=0.65]{../diagrams/database-sharding-by-region}
  \caption{Example of Data-Centric Storage for Sensor Networks using Database
  Sharding}
  \label{fig:database-sharding-by-region}
\end{figure}

%This data organization architecture is currently being used to manage
%structured data that scales to petabytes or more data. One example of such use
%is on BigTable \cite{bigtable}, which provides support to different Google
%products with requirements on high availability and dynamic data models and
%uses automatic sharding to its data. The BigTable project uses MapReduce
%[ds02], a programming model designed to search huge amounts of data in
%paralell.

@@

\begin{itemize}
  \item Referential integrity: any cross-shard queries can represent problems 
  since the application layer is responsible to enforce data integrity, like
  verifying constraints of foreign keys;
  \item Rebalancing Shards: If the schema model changes for the shards, the
  strategy of rebalancing must be done. For instance, when one of the shards 
  outgrows more than the others. As a consequence, the database shards must be
  rebalanced by transferring data to new locations whenever necessary. This
  technique is starting to appear as COTS implemetations such as
  \cite{mongodb}.
\end{itemize}

@@

Other projects have been using the XML data models. It falls into the same
category as the Relational Model, since the XML documents \cite{xml}, when in
the context of a database, must be complaint to an XML Schema \cite{xml-schema}. As shown in
section 3, this model uses XPath \cite{xml-xpath} technology for querying
documents, although some hibrid technologies may still use SQL \cite{sql} for
that matter. In this way, an instance of such schema dependent types can be
seen in figure \ref{fig:persistence-example-relational}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.7]{../diagrams/persistence-example-relational}
  \caption{Instance of a Relational Model Database Prototype}
  \label{fig:persistence-example-relational}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.7]{../diagrams/KVP-on-Relational-Model}
  \caption{KVP Data Model implementation using Relational Model}
  \label{fig:KVP-on-Relational-Model}
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.7]{../diagrams/persistence-example-relational-kvp}
  \caption{KVP over Relational Model Database Prototype}
  \label{fig:persistence-example-relational-kvp}
\end{figure}

@@

Sensor Network data can be saved in either Centralized or Distributed Systems.
As described in  While Centralized Database Systems are easier to manage, it may
face challenges regarding its data. For this reason, the use of partitioning or
sharding would be an addition to persist sensor data.
